---
title: "New York Educational Data"
output: html_notebook
---

This notebook will analyze the New York Grauduation Data located at 
https://data.nysed.gov/downloads.php

The majority of the analysis will use SQLite for the SQL version

```{r}
# Import dataset
library(readr)
f = "C:/Users/yaleq/Google Drive/SQL Project/GRAD_RATE_AND_OUTCOMES_2019.csv"
grad_df <- read_csv(f)
```

```{r message=FALSE}
# Import needed libraries
library(sqldf)
library(tidyverse)
library(dplyr)
```

```{r}
# Count how many rows are in the dataset

query = "SELECT COUNT(1) AS n_rows
         FROM grad_df"
sqldf(query)

```

```{r}
# Column Names and information
query = "PRAGMA table_info(grad_df)"

sqldf(query)
```

```{r}
# Check how to aggregate the data.
query = "SELECT DISTINCT aggregation_type
         FROM grad_df
         ORDER BY aggregation_type DESC"

sqldf(query)
```

```{r}
# Number of unique schools
query = "SELECT COUNT (DISTINCT aggregation_name) as 'num_schools'
         FROM grad_df
         WHERE aggregation_type = 'School'"
         

sqldf(query)
```

Schools are reporting based on subgroups.  I want to see what subgroups
are reported

```{r}
query = "SELECT DISTINCT subgroup_name
         FROM grad_df
         WHERE aggregation_type = 'School'
         ORDER BY subgroup_name"
sqldf(query)

```

```{r}
# Average graduation percentage
query = "SELECT subgroup_name, AVG(grad_pct) as 'avg_grad_percent'
         FROM grad_df
         WHERE aggregation_type = 'School'
         GROUP BY subgroup_name
         ORDER BY AVG(grad_pct) DESC"

sqldf(query)
```

From the above table we can see that female, and students who are not
economically disadvantaged, on average, graduate at a higher percentage
than the overall graduation percentage.

It looks like schools report graduation rate based on cohort and years
enrolled

```{r}
# Cohorts
query = "SELECT DISTINCT membership_desc
         FROM grad_df
         WHERE aggregation_type = 'School'
         ORDER BY membership_desc"

sqldf(query)
```



```{r}
# 4-year grad rate for high schools
query = "SELECT aggregation_name AS 'School Name', 
                membership_desc AS 'Student Type',
                grad_pct AS 'Grad Percentage'
         FROM grad_df
         WHERE aggregation_type = 'School' 
               AND subgroup_name = 'All Students'
               AND aggregation_name LIKE '%high%'
               AND membership_desc LIKE '%4 Year Outcome - August 2019%'
               AND grad_pct >= 0
         GROUP BY aggregation_name
         ORDER BY grad_pct DESC
        "

sqldf(query)
```

Looking at charter vs non-charter high schools
```{r}
# 4-year grad rate comparison of all students
query = "SELECT nrc_desc AS 'School Type', 
                membership_desc AS 'Student Type',
                grad_pct AS 'Grad Percentage'
         FROM grad_df
         WHERE subgroup_name = 'All Students'
               AND aggregation_name LIKE '%high%'
               AND membership_desc LIKE '%4 Year Outcome - August 2019%'
               AND grad_pct >= 0
         GROUP BY nrc_desc
         ORDER BY grad_pct DESC
        "

sqldf(query)
```

Let's look at dropout rates
```{r}
# dropout grad rate comparison
query = "SELECT nrc_desc AS 'School Type', 
                membership_desc AS 'Student Type',
                ROUND(AVG(dropout_pct),2) AS 'Average Dropout Percentage'
         FROM grad_df
         WHERE subgroup_name = 'All Students'
               AND aggregation_name LIKE '%high%'
               AND membership_desc LIKE '%4 Year Outcome - August 2019%'
               AND grad_pct >= 0
         GROUP BY nrc_desc
         ORDER BY dropout_pct DESC
        "

sqldf(query)
```















